# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r loaddata, echo=TRUE}
library(ggplot2)
rawdata <- read.csv("~/Documents/coursera//R_computing/DS spec//Reproducible Research//RepData_PeerAssessment1//activity.csv", sep=',', na.strings="NA", header=T, colClasses = c("numeric", "Date", "numeric"))
rawdata$day<-weekdays(rawdata$date)
```


## What is mean total number of steps taken per day?
Histogram of total number of steps taken each day
```{r stepsperday, echo=TRUE}
steps_per_day<-tapply(rawdata$steps, rawdata$date, sum)
stepsperday<-unlist(steps_per_day)
qplot(stepsperday)
```

Median of total number of steps taken each day
```{r medianofsteps, echo=TRUE}
# Mean
mean(stepsperday, na.rm=T)
# Median
median(stepsperday, na.rm=T)
```

## What is the average daily activity pattern?
Average number of steps made per interval

```{r stepsperinterval, echo=TRUE}
stepsperinterval <- aggregate(steps ~ interval, data=rawdata, mean, na.rm = TRUE)
ggplot(stepsperinterval, aes(x=interval,y=steps)) + geom_line(color="red", size=1)
```


And the most active 5-minute interval is
```{r mostactive, echo=TRUE}
stepsperinterval[which.max(stepsperinterval$steps),]$interval
```
## Imputing missing values

```{r nanumber, echo=TRUE}
sum(is.na(rawdata))
```

The method is simple: assuming that in different weekdays we walk different amount of steps
We imput for each 5-min interval the average number of steps per such interval in a concrete day of week

```{r inputmissing, echo=TRUE}
avstperday <- aggregate(steps ~ interval + day, rawdata, mean, na.rm = TRUE)
data <- merge(rawdata, avstperday, by=c("interval", "day"))
data <- transform(data, steps.x = ifelse(is.na(steps.x),steps.y,steps.x))
data <- data.frame(data[,1:4])
names(data) <- c("interval", "day","steps", "date")
data$steps <- round(data$steps, digits = 0)
data <- data[order(data$date, data$interval),]
```

Histogram of total number of steps taken each day

```{r histoftotal, echo=TRUE}
datasplitbyday <- aggregate(steps ~ date, data, sum, na.rm = TRUE)
qplot(steps, data = datasplitbyday)
```

Mean and median for new data
```{r meanandmedian, echo=TRUE}
#Mean
mean(datasplitbyday$steps)

#Median
median(datasplitbyday$steps)
```


